Connecting the Popularity Adjusted Block Model to the Generalized Random Dot Product Graph for Community Detection and Parameter Estimation

### 2021-03-02 summary.Rmd
fixed proof of theorem 5

**Definition 4** (Inradius) \cite{soltanolkotabi2012} 
\cite{pmlr-v28-wang13}. 
The inradius of a convex body $\mathcal{P}$, denoted by $r(\mathcal{P})$, is 
defined as the radius of the largest Euclidean ball inscribed in $\mathcal{P}$.

**Lemma 5.1**. 
Let $P_n = V_n \Lambda_n V_n^\top$ be defined as in Lemma 5.1. Let 
$A_n \sim \Bernoulli(P_n)$ be the adjacency matrix drawn from $P_n$, and let 
$A_n = \hat{V}_n \hat{\Lambda}_n \hat{V}_n^\top$ be its approximate spectral
decomposition using the $K (K+1) / 2$ most positive and $K (K-1) / 2$ most 
negative eigenvalues and their corresponding eigenvectors.
Without loss of generality, order the rows of $V_n$ and $\hat{V}_n$ by 
community, and let $V_n^{(k)}$ and $\hat{V}_n^{(k)}$ be the $n_k \times K^2$ 
matrix of the rows of the $k$^th^ community in $V_n$ and $\hat{V}_n$ 
respectively. 
Let $\mu_n^{(k)} = \mu(V_n^{(k)})$ be the subspace incoherence of $V_n^{(k)}$ 
as defined by \citet{soltanolkotabi2012}. Then

@@

\begin{equation} \label{eq:mu-conv}
\hat{\mu}_n^{(k)} \stackrel{a.s.}{\to} 0
\end{equation}

@@

For the second part, 
this falls directly out of Theorem 5 of \citet{rubindelanchy2017statistical}. 
Since $\hat{V}_n^{(k)} \stackrel{a.s.}{\to} V_n^{(k)}$ and $\mu(\cdot)$ is 
continuous, $\mu(\hat{V}_n^{(k)}) \stackrel{a.s.}{\to} \mu(V_n^{(k)}) = 0$.

@@

\begin{equation} \label{eq:thm5-proof}
r(\hat{V}_n^{(k)}) - \mu(\hat{V}_n^{(k)}) \stackrel{a.s.}{\to} 
r(\hat{V}_n^{(k)}) > 0
\end{equation}

for each $k = 1, ..., K$.

Then Theorem 6 of \citet{pmlr-v28-wang13}, (\ref{eq:delta-conv}) and
(\ref{eq:thm5-proof}) imply that the SSC detection property holds as 
$n \to infty$ with probability 1.

>>>>

**Proof of Theorem 4**.

**Definition 4** (Inradius) \cite{soltanolkotabi2012} 
\cite{pmlr-v28-wang13}. 
The inradius of a convex body $\mathcal{P}$, denoted by $r(\mathcal{P})$, is 
defined as the radius of the largest Euclidean ball inscribed in $\mathcal{P}$.
In addition, $r(X)$ for data matrix $X$ with rows $x_i^\top$ represents 
the inradius of the symmetric convex hull of $X$.

**Lemma 5.1**. 
Let $P_n = V_n \Lambda_n V_n^\top$ be the edge probability matrix of a PABM 
with $n$ vertices. Let $A_n \sim \Bernoulli(P_n)$ be the adjacency matrix 
drawn from $P_n$, and let $A_n = \hat{V}_n \hat{\Lambda}_n \hat{V}_n^\top$ be 
its approximate spectral decomposition using the $K (K+1) / 2$ most positive 
and $K (K-1) / 2$ most negative eigenvalues and their corresponding 
eigenvectors.Without loss of generality, order the rows of $V_n$ and 
$\hat{V}_n$ by community, and let $V_n^{(k)}$ and $\hat{V}_n^{(k)}$ be the 
$n_k \times K^2$ matrix of the rows of the $k$^th^ community in $V_n$ and 
$\hat{V}_n$ respectively. 
Let $\mu_n^{(k)} = \mu(V_n^{(k)})$ be the subspace incoherence of $V_n^{(k)}$ 
as defined by \citet{soltanolkotabi2012}. Then

@@

\begin{equation} \label{eq:mu-conv}
\hat{\mu}_n^{(k)} = 0
\end{equation}

@@

For the second part, 
$\mu(\hat{V}_n^{(k)}) = \max_{v_n^{(i)} \in V_n^{(-k)}} ||U_n^{(k)} v_n^{(i)}||_\infty$ where the rows of $U_n^{(k)}$ consist of vectors lying on the subspace spanned by $V_n^{(k)}$. Since if $u_n^{(i)} \in \text{span}(V_n^{(k)}) \implies u_n^{(i)} \perp v_n^{(j)}$ $\forall v_n^{(j)} \in V_n^{(-k)}$ by Theorem 2, this quantity must be 0.

@@

\begin{equation} \label{eq:thm5-proof}
r(V_n^{(k)}) - \mu(\hat{V}_n^{(k)}) = r(V_n^{(k)})
\end{equation}

for each $k = 1, ..., K$. If enough points are drawn from each community that 
the corresponding uncorrupted embedding points span a $K$-dimensional subspace
(Theorem 2), then $r(V_n^{(k)}) > 0$.

Then by Theorem 6 of \citet{pmlr-v28-wang13}, (\ref{eq:delta-conv}) and
(\ref{eq:thm5-proof}) imply that the SSC detection property holds as 
$n \to \infty$ with probability 1.
